---
title: "Run an external command from R and record its output."
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Run an external command from R

The `ngsmisc` package provides some wrapper functions for easy handling of `processx::run()`.

The procedure for executing a command line statement is as follows.

1. Split the command line statement with `tokenize_cmd()`.
2. Pass the split statement to `run_cmd()` and execute it.
3. Check the execution results of output.

To run a external command line statement using `processx::run()`, 
you have to split the statement into the command and each argument.
You can separate the command line statement by white spaces using `tokenize_cmd()`.

```{r}
# Define an example command line statement.
cmd <- "echo 'This is stdout.'"

# Split the statement.
# The first element is the name of command you want to run.
# And rests are the arguments of it.
ngsmisc::tokenize_cmd(cmd)
```

The splitted statement can be executed using `run_cmd()`.
The `run_cmd()` passes the first element of the first argument to the `command` 
of `processx::run()` and the rest of the first argument to the `args`.
The rest of the arguments are passed directly to `processx::run()`.

The `run_cmd()`(`processx::run()`) invisibly returns the result of execution as the named list.

```{r}
# Run the separated command statement using `run_cmd()`.
out <-
  ngsmisc::tokenize_cmd(cmd) |>
  ngsmisc::run_cmd()

# The output is a named list containing, execution status, 
# output of stdout, output of stderr, and whether timeout or not.
out
```

The output is a list, so you can easily print and check the results in it.
The `ngsmisc` package also provides some deadly simple wrapper functions to get/print the results.

```{r}
# Write a simple shell script, which outputs both stdout and stderr.
c(
  "echo 'This is stdout.' >&1",
  "echo 'This is stderr.' >&2"
) |>
  readr::write_lines("temp.sh")

# Execute above shell script and get results.
out <- ngsmisc::tokenize_cmd("bash temp.sh") |> ngsmisc::run_cmd()

# Get the stdout result.
out$stdout
```

```{r, echo=FALSE}
fs::file_delete("temp.sh")
```

# Shortened form

The `ngsmisc` also provides several shortened functions that performs the above 
processes in a single function.

```{r}
# Do the same thing above by `run_get_stdout()`
ngsmisc::run_get_stdout(cmd)
```

# Save the result of command line execution

Sometimes I think it would be useful to save the results in a file so I can look at them later.
This is because if the results are saved (cached), there is no need to run it again.
It is especially true when you run a process that takes a long time.

When saving the results to a text file, you can keep the directory clean if you can save the files in one place.
The `path_cmdout()` is a convenience function that returns the file path to save the output,  and also makes the destination directory.

```{r}
# `path_cmdout()` inserts specified directory by the `save_dir` argument 
# (default: `"cmdout_cache"`) in between the directory specified by the `wd` and
# file path (`...`).
# if `create_dir` is `TRUE`, intermediate directories will be created.
args(ngsmisc::path_cmdout)

ngsmisc::path_cmdout("/path/to/workdir", "cmdout.txt", create_dir = FALSE)

# If you are working with RStudio's project feature,
# you can use `here::here()` as follows.
fpath <- ngsmisc::path_cmdout(here::here(), "cmdout.txt")

# Save the output to a file, and then see the result later as follows.
ngsmisc::run_get_stdout(cmd) |> readr::write_lines(file = fpath)
readr::read_lines(fpath) |> cat(sep = "\n")
```

```{r, echo=FALSE}
fs::file_delete(fpath)
fs::file_delete(fs::path_dir(fpath))
```

# Sessioninfo
```{r}
sessioninfo::package_info()
```
