---
title: "Run an external command from R and record its output."
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Run an external command from R

The `ngsmisc` package provides wrapper functions for handling `processx::run()`.

The typical procedure for executing a command line statement in `ngsmisc` is as follows.

1. Split the command line statement using `tokenize_cmd()`.
2. Pass the split statement to `run_cmd()` and execute it.
3. Check the execution results.

To run a external command line statement with `processx::run()`, 
you have to split the statement into the command and each argument.
You can do this easily with `tokenize_cmd()`, which separates the statement by whitespace.

```{r}
# Define an example command line statement.
cmd <- "echo 'This is stdout.'"

# Split the statement.
# The first element is the command to run.
# The remaining elements are its arguments.
ngsmisc::tokenize_cmd(cmd)
```

The splitted statement can then be executed using `run_cmd()`.
`run_cmd()` passes the first element of the first argument to the `command` of `processx::run()`,
and the remaining elements to the `args` parameter.
Any additional arguments are passed to `processx::run()`.

`run_cmd()` (like `processx::run()`) invisibly returns the execution result as a named list.

```{r}
# Run the separated command statement using `run_cmd()`.
out <-
  ngsmisc::tokenize_cmd(cmd) |>
  ngsmisc::run_cmd()

# The output is a named list containing execution status, 
# stdout, stderr, and whether timeout or not.
out
```

Since the output is a list, you can easily print and check the results.
The `ngsmisc` package also provides simple wrapper functions to get/print the results.

```{r}
# Write a simple shell script that outputs both stdout and stderr.
c(
  "echo 'This is stdout.' >&1",
  "echo 'This is stderr.' >&2"
) |>
  writeLines("temp.sh")

# Execute the shell script and get the results.
out <- ngsmisc::tokenize_cmd("bash temp.sh") |> ngsmisc::run_cmd()

# Get the stdout result.
out$stdout
```

```{r, echo=FALSE}
fs::file_delete("temp.sh")
```

# Shortened form

`ngsmisc` also provides several shortcut functions that performs the above processes in a single step.

```{r}
# Do the same thing above by `run_get_stdout()`
ngsmisc::run_get_stdout(cmd)
```

# Saving the result of command execution

It could be useful to save the results to a file for later reference.
If the results are saved (cached), you don't need to rerun the command,
which is especiall helpful for long-running processes.

When saving results to a text file, it's convienient to keeep all output files in one directory.
The `path_cmdout()` function returns a file path for saving the output and creates the destination directory if needed.

```{r}
# `path_cmdout()` inserts the directory specified by `save_dir` (default: `"cmdout_cache"`)
# between the working directory (`wd`) and the fspecified by the file path (`...`).
# if `create_dir` is `TRUE`, intermediate directories are created.
args(ngsmisc::path_cmdout)

ngsmisc::path_cmdout("/path/to/workdir", "cmdout.txt", create_dir = FALSE)

# If you are using RStudio's project feature,
# you can use `here::here()` as follows.
fpath <- ngsmisc::path_cmdout(here::here(), "cmdout.txt")

# Save the output to a file, then you can see the result later.
ngsmisc::run_get_stdout(cmd) |> readr::write_lines(file = fpath)
readLines(fpath) |> cat(sep = "\n")
```

```{r, echo=FALSE}
fs::file_delete(fpath)
fs::file_delete(fs::path_dir(fpath))
```

# Sessioninfo
```{r}
sessioninfo::package_info()
```
